---
title: "Data Visualization in `rrr`"
author: "Chris Addy"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## The Philosophy of Graphics in `rrr`

The graphical display of data should be intuitive enough that any viewer of the display unfamiliar with the data, even unfamiliar with statistical analysis generally, should be able to see clear patterns, if in fact clear patterns exist in the data. This, the author believes, is especially true for multivariate data, where due to the higher dimensionality of the data, the potential for visualization to overwhelm is high.

```{r}
library(dplyr)

### LOAD DATA

### TOBACCO DATA SET
data(tobacco)

tobacco_x <- tobacco %>%
	select(starts_with("X"))

tobacco_y <- tobacco %>% 
	select(starts_with("Y"))

### PENDIGITS DATA SET
data(pendigits)
digits <- as_data_frame(pendigits)

digits_class <- digits %>% select(V35)
digits_features <- digits %>% select(-V35, -V36)

### COMBO-17 GALAXY DATA SET
data(COMBO17)
galaxy <- as_data_frame(COMBO17) %>%
       select(-starts_with("e."), -Nr, -UFS:-IFD) %>%
       na.omit()

### IRIS DATA SET
data(iris)
iris <- as_data_frame(iris)

iris_features <- iris %>%
	select(-Species)

iris_class <- iris %>%
	select(Species)

### COMBO-17 DATA SET
data(COMBO17)
galaxy <- as_data_frame(COMBO17) %>%
       select(-starts_with("e."), -Nr, -UFS:-IFD) %>%
       na.omit()

galaxy_x <- galaxy %>%
	select(-Rmag:-chi2red)

galaxy_y <- galaxy %>%
	select(Rmag:chi2red)
```

```{r}
pca_rank_trace(digits_features)

pca_pairwise_plot(digits_features)

pca_pairwise_plot(digits_features, pc_x = 1, pc_y = 3, class_labels = digits_class)
```

## Rank Trace Plots

```{r}
gamma1 <- diag(1, 3)
gamma2 <- solve(cov(tobacco_y))

rank_trace(tobacco_x, tobacco_y, gamma1)

rank_trace(tobacco_x, tobacco_y, gamma2)
```

Rank trace plots can be made interactive with the argument `interactive = TRUE`.

```{r}
rank_trace(tobacco_x, tobacco_y, gamma1, interactive = TRUE)

rank_trace(tobacco_x, tobacco_y, gamma2, interactive = TRUE)
```

## RRR Residuals Plots

```{r}
#rrr_residuals_plot(tobacco_x, tobacco_y, ident, rank = 1)
```

## PC Rank Trace Plots

```{r}
pca_rank_trace(digits_features)
```

PCA rank trace plots can be made interactive with the argument `interactive = TRUE`.

```{r}
pca_rank_trace(digits_features, interactive = TRUE)
```


## PC Pairwise Plots

```{r}
pca_pairwise_plot(digits_features, pc_x = 1, pc_y = 3, class_labels = digits_class, interactive = TRUE)
```

## PCA 3D Plots

```{r}
pca_3D_plot(digits_features, class_labels = digits_class)
```

### Change Point Size with argument `point_size`

```{r}
pca_3D_plot(iris_features, class_labels = iris_class)
pca_3D_plot(iris_features, class_labels = iris_class, point_size = 5)
```

## CVA Rank Trace Plots

```{r}
cva_rank_trace(galaxy_x, galaxy_y)
```

```{r}
cva_rank_trace(galaxy_x, galaxy_y, interactive = TRUE)
```

## CVA Pairwise Plots

```{r}
cva_pairwise_plot(galaxy_x, galaxy_y, cv_pair = 2)
```

```{r}
cva_pairwise_plot(galaxy_x, galaxy_y, cv_pair = 2, interactive = TRUE)
```

## CVA Residuals Plots

```{r}
cva_residual_plot(galaxy_x, galaxy_y, rank = 2)
```

## CVA 3D Residuals Plots

```{r}
cva_residual_3D_plot(galaxy_x, galaxy_y)
```

## LDA Pairwise Plots

```{r}
lda_pairwise_plot(iris_features, iris_class)
```

```{r}
lda_pairwise_plot(iris_features, iris_class, interactive = TRUE)
```

## LDA 3D Plots

```{r}
args(lda_3D_plot)
```

```{r}
lda_3D_plot(digits_features, digits_class, k = 0.0001)
```