---
title: "Data Visualization in rrr"
author: "Chris Addy"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## The Philosophy of Graphics in `rrr`

The graphical display of data should be intuitive enough that any viewer of the display unfamiliar with the data, even unfamiliar with statistical analysis generally, should be able to see clear patterns, if in fact clear patterns exist in the data. This, the author believes, is especially true for multivariate data, where due to the higher dimensionality of the data, the potential for visualization to overwhelm is high.

```{r}
library(dplyr)

### LOAD DATA

### TOBACCO DATA SET
data(tobacco)

tobacco_x <- tobacco %>%
	select(starts_with("X"))

tobacco_y <- tobacco %>% 
	select(starts_with("Y"))

### PENDIGITS DATA SET
data(pendigits)
digits <- as_data_frame(pendigits)

digits_class <- digits %>% select(V35)
digits_features <- digits %>% select(-V35, -V36)

### COMBO-17 GALAXY DATA SET
data(COMBO17)
galaxy <- as_data_frame(COMBO17) %>%
       select(-starts_with("e."), -Nr, -UFS:-IFD) %>%
       na.omit()

### IRIS DATA SET
data(iris)
iris <- as_data_frame(iris)

iris_features <- iris %>%
	select(-Species)

iris_class <- iris %>%
	select(Species)

### COMBO-17 DATA SET
data(COMBO17)
galaxy <- as_data_frame(COMBO17) %>%
       select(-starts_with("e."), -Nr, -UFS:-IFD) %>%
       na.omit()

galaxy_x <- galaxy %>%
	select(-Rmag:-chi2red)

galaxy_y <- galaxy %>%
	select(Rmag:chi2red)
```

## Rank Trace Plots

`rank_trace()` can draw rank trace plots for reduced-rank regression, principal components analysis and canonical variate analysis by setting `type = "identity"` (the default), `type = "pca"`, or `type = "cva"`, respectively.

```{r}
rank_trace(tobacco_x, tobacco_y)
rank_trace(digits_features, digits_features, type = "pca")
rank_trace(galaxy_x, galaxy_y, type = "cva")
```

Rank trace plots can be made interactive with the argument `interactive = TRUE`.

```{r}
rank_trace(tobacco_x, tobacco_y, interactive = TRUE)
rank_trace(digits_features, digits_features, type = "pca", interactive = TRUE)
rank_trace(galaxy_x, galaxy_y, type = "cva", interactive = TRUE)
```

## Residual Plots

```{r}
residuals(tobacco_x, tobacco_y, rank = 1)
residuals(galaxy_x, galaxy_y, type = "cva")
```

## Pairwise Plots

Pairwise plots can be created for PCA, CVA, LDA, by setting `type = "pca"`, `type = "cva"`, and `type = "lda"`, respectively.

```{r}
pairwise_plot(digits_features, digits_class, pair_x = 1, pair_y = 3)

pairwise_plot(galaxy_x, galaxy_y, type = "cva", pair_x = 2)

pairwise_plot(iris_features, iris_class)
```

Pairwise plots can be turned into interactive plotly graphs by setting `interactive = TRUE`.

pairwise_plot(digits_features, digits_class, pair_x = 1, pair_y = 3, interactive = TRUE)

pairwise_plot(galaxy_x, galaxy_y, pair_x = 2, interactive = TRUE)

pairwise_plot(iris_features, iris_class, interactive = TRUE)
```

## 3D Plots

```{r}
threewise_plot(digits_features, digits_class, type = "pca")
```

```{r}
threewise_plot(galaxy_x, galaxy_y, type = "cva")
```

```{r}
threewise_plot(digits_features, digits_class, type = "lda", k = 0.001)
```

### Change Point Size with argument `point_size`

```{r}
threewise_plot(iris_features, iris_class, type = "pca")
threewise_plot(iris_features, iris_class, point_size = 5)
```