---
title: "`rrr` for Linear Discriminant Analysis"
author: "Chris Addy"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Mathematical Background

### LDA is a Special Case of [CVA](c-cva.html)

Linear discriminant analysis is a classification procedure. We can turn it into a regression procedure -- specifically a reduced-rank canonical variate procedure -- in the following way.

Let each $i = 1, 2, \dots, n$ observation observationbelong to one, and only one, of $K = s + 1$ distinct classes.

We can construct an *indicator response matrix*, $\mathbf{Y}$

$$
\mathbf{Y}
$$

```{r}
### create training and test sets
set.seed(12345)
num_obs <- dim(digits)[1]
train_number <- floor(.75 * num_obs)
train_index <- sample(num_obs, train_number)

### training set
train_digits_features <- digits_features[train_index, ]
train_digits_class <- digits_class[train_index, ]

### test set
test_digits_features <- digits_features[-train_index, ]
test_digits_class <- digits_class[-train_index, ]
```

## Assesss Effective Dimensionality

## Diagnostics

## Fit LDA Model

### Fit LDA model with `lda()`

```{r}
lda(digits_features, digits_class, rank = 3)
```