---
title: "`rrr` for Linear Discriminant Analysis"
author: "Chris Addy"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Mathematical Background

### LDA is a Special Case of [CVA](cva.html)

Linear discriminant analysis is a classification procedure. We can turn it into a regression procedure -- specifically a reduced-rank canonical variate procedure -- in the following way.

Let each $i = 1, 2, \dots, n$ observation observationbelong to one, and only one, of $K = s + 1$ distinct classes.

We can construct an *indicator response matrix*, $\mathbf{Y}$

$$
\mathbf{Y}
$$

## The `iris` Data Set

```{r}
library(dplyr)
data(iris)
iris <- as_data_frame(iris)

glimpse(iris)
```

```{r}
iris_features <- iris %>%
	select(-Species)

iris_class <- iris %>%
	select(Species)
```

## Assesssing Effective Dimensionality

## Fitting LDA Models

### Fit LDA model with `lda()`

```{r}
lda(iris_features, iris_class, k = 0.0001)
```

### Calculate LDA Scores with `lda_scores()`

```{r}
lda_scores(iris_features, iris_class, k = 0.0001)
```

## Diagnostics

## Plotting

### Plot LDA Pairs with `lda_pairwise_plot()`

```{r}
args(lda_pairwise_plot)
```

A typical graphical display for multiclass LDA is to plot the $j$th discriminant scores for the $n$ points against the $k$ discriminant scores with the class means superimposed on the same graph.

```{r}
lda_pairwise_plot(iris_features, iris_class, k = 0.0001)
```