---
title: "`rrr` for Multivariate Analysis"
author: "Chris Addy"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r message = FALSE}
devtools::install_github("chrisaddy/rrr")
library(rrr)
```

```{r}
library(dplyr)
```

$$ X_1 + Y = 2 $$

### Reduced-Rank Regression and Its Special Cases

The multivariate linear regression model
```{r}
library(MMST)
library(gridExtra)
```

$$
\overset{s \times 1}{\mathbf{Y}} = \overset{s \times 1}{\boldsymbol{\mu}} + \overset{s \times r}{\mathbf{C}} \overset{r \times 1}{\mathbf{X}} + \overset{s \times 1}{\varepsilon}
$$

## Principal Components Analysis

Set

$$latex
\begin{aligned*}
	\mathbf{Y} & \equiv \mathbf{X} \\
	s & = r \\
	\mathbf{\Gamma} & = \mathbf{I}_r \\
\end{aligned*}
$$
	
Then, the least-squares error criterion

\begin{equation}
	\begin{aligned}
	\mathrm{E}\left[\left(\mathbf{X} - \boldsymbol{\mu} - \mathbf{A}\mathbf{B}\mathbf{X}\right)^\tau\left(\mathbf{X} - \boldsymbol{\mu} - \mathbf{A}\mathbf{B}\mathbf{X}\right)\right]
	\end{aligned}
\end{equation}

is minimized by the reduced-rank regression solution,

\begin{equation}
	\begin{aligned}
		\mathbf{A}^{\left(t\right)} & = \left(\mathbf{v}_1, \dots, \mathbf{v}_t\right) \\
		\mathbf{B}^{\left(t\right)} & = \mathbf{A}^{\left(t\right) \tau} \\
		\boldsymbol{\mu}^{\left(t\right)} & = \left(\mathbf{I}_r - \mathbf{A}^{\left(t\right)}\mathbf{B}^{\left(t\right)}\right)\boldsymbol{\mu}_X \\
	\end{aligned}
\end{equation}

where $\mathbf{v}_j = \mathbf{v}_j \left(\mathbf{\Sigma}_{XX}\right)$ is eigenvector associated with the $j$th largest eigenvalue of  $\mathbf{\Sigma}_{XX}.$

The best reduced-rank approximation to the original $\mathbf{X}$ is

\begin{equation}
	\begin{aligned}
		\hat{\mathbf{X}}^{\left(t\right)} & =
		\boldsymbol{\mu}^{\left(t\right)} + \mathbf{A}^{\left(t\right)}\mathbf{B}^{\left(t\right)} \mathbf{X}
	\end{aligned}
\end{equation}

```{r}
data(pendigits)

digits <- as_data_frame(pendigits)
digits_class <- digits %>% select(V35)
digits_features <- digits %>% select(-V35, -V36)

#pca_rank_trace(digits_features)

#pca_rank_trace_plot(digits_features)

#pc_pairwise_plot(digits_features, class_labels = digits_class)

#pc_plot_3D(digits_features, class_labels = digits_class)
```

## Canonical Variate Analysis

$\mathbf{X}, \mathbf{Y}$ are jointly distributed, with

\begin{equation}
	\begin{aligned}
		\mathrm{E}\left\{
		\begin{pmatrix}
			\mathbf{X} \\
			\mathbf{Y} \\
		\end{pmatrix}
		\right\} =
		\begin{pmatrix}
			\boldsymbol{\mu}_X \\
			\boldsymbol{\mu}_Y \\
		\end{pmatrix}, \quad
		\mathrm{cov}\left\{
		\begin{pmatrix}
			\mathbf{X} \\
			\mathbf{Y} \\
		\end{pmatrix}\right\} =
		\begin{pmatrix}
			\mathbf{\Sigma}_{XX} & \mathbf{\Sigma}_{XY} \\
			\mathbf{\Sigma}_{YX} & \mathbf{\Sigma}_{YY} \\
		\end{pmatrix}
	\end{aligned}
\end{equation}

Let $1 \leq t \leq \mathrm{min}\left(r, s\right)$. $\mathbf{X}$ and $\mathbf{Y}$ are linearly projected into new vector variates,

\begin{equation}
	\begin{aligned}
		\boldsymbol{\xi}_{t \times 1} = \mathbf{G}_{t \times r} \mathbf{X}_{r \times 1}, \quad \boldsymbol{\omega}_{t \times 1} = \mathbf{H}_{t \times s} \mathbf{X}_{s \times 1}
		\end{aligned}
		\end{equation}

Finding a solution to the problem

\begin{equation}
	\begin{aligned}
		\boldsymbol{\xi} \approx \boldsymbol{\nu} + \boldsymbol{\omega}
	\end{aligned}
\end{equation}

requires minimizing the least-squares criterion

\begin{equation}
	\begin{aligned}
		\mathrm{E}\left[\left(\mathbf{H}\mathbf{Y} - \boldsymbol{\nu} - \mathbf{G}\mathbf{X}\right)\left(\mathbf{H}\mathbf{Y} - \boldsymbol{\nu} - \mathbf{G}\mathbf{X}\right)^\tau\right]
	\end{aligned},

where we assume $\mathrm{cov}\left(\boldsymbol{\omega}\right) = \mathbf{I}_t$.
\end{equation}. This least-squares criterion is minimized when

\begin{equation}
	\begin{aligned}
		\boldsymbol{\nu}^{\left(t\right)} & = \mathbf{H}^{\left(t\right)} \boldsymbol{\mu}_Y - \mathbf{G}^{\left(t\right)}\boldsymbol{\mu}_Y \\
		\mathbf{G}^{\left(t\right)} & =
		\begin{pmatrix}
			\mathbf{v}_1^\tau \\
			\vdots \\
			\mathbf{v}_t^\tau \\
		\end{pmatrix}
		\mathbf{\Sigma}_{YY}^{-1/2} \mathbf{\Sigma}_{YX}\mathbf{\Sigma}_{XX}^{-1} \\
		\mathbf{H}^{\left(t\right)} & =
		\begin{pmatrix}
			\mathbf{v}_1^\tau \\
			\vdots \\
			\mathbf{v}_t^\tau \\
		\end{pmatrix}
		\mathbf{\Sigma}_{YY}^{-1/2} \\
	\end{aligned}
\end{equation}

where $\mathbf{v}_j$ is the eigenvector associated with the $j$th largest eigenvalue of

\begin{equation}
	\begin{aligned}
		\mathbf{R} & = \mathbf{\Sigma}_{YY}^{-1/2}\mathbf{\Sigma}_{YX} \mathbf{\Sigma}_{XX}^{-1} \mathbf{\Sigma}_{XY} \mathbf{\Sigma}_{YY}^{-1/2}
	\end{aligned}
\end{equation}

```{r}
### COMBO-17 galaxy data
data(COMBO17)
galaxy <- as_data_frame(COMBO17) %>%
       select(-starts_with("e."), -Nr, -UFS:-IFD) %>%
       na.omit()

galaxy_x <- galaxy %>% select(-Rmag:-chi2red)
galaxy_y <- galaxy %>% select(Rmag:chi2red)

#galaxy_x
#galaxy_y
```

## Linear Discriminant Analysis


Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
